---
title: "Applications"
author: "Marc Weber"
output: 
  html_document:
    theme: flatly
    keep_md: yes
    number_sections: true
    highlighted: default 
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: no
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Applications withs StreamCatTools}
  %\VignetteEncoding{UTF-8}{inputenc}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  warning = FALSE, 
  message = FALSE
)
```

### Plotting watershed data
In this example we access a single variable for the Calapooia River using `sc_get_data` function. We then use the `nhdplusTools` library to grab flowlines and watershed for the Calapooia, plot the selected StreamCat metric for the Calapooia River and show the watershed.
```{r wshd, results='hide'}
library(StreamCatTools)
start_comid = 23763529
nldi_feature <- list(featureSource = "comid", featureID = start_comid)

flowline_nldi <- nhdplusTools::navigate_nldi(nldi_feature, mode = "UT", data_source = "flowlines", distance=5000)

# get StreamCat metrics
temp_dir <- '.'
nhdplus <- nhdplusTools::subset_nhdplus(comids = as.integer(flowline_nldi$UT_flowlines$nhdplus_comid), output_file = file.path(temp_dir, "nhdplus.gpkg"),nhdplus_data = "download",overwrite = TRUE,flowline_only = FALSE)

names(nhdplus)
cats <- nhdplus$CatchmentSP
comids <- paste(cats$featureid,collapse=",",sep="")

df <- sc_get_data(metric='PctImp2011', aoi='catchment', comid=comids)

flowline_nldi <- flowline_nldi$UT_flowlines
flowline_nldi$PCTIMP2011CAT <- df$PCTIMP2011CAT[match(flowline_nldi$nhdplus_comid, df$COMID)]

basin <- nhdplusTools::get_nldi_basin(nldi_feature = nldi_feature)
```

```{r wshd pt2}
library(mapview)
mapview::mapviewOptions(fgb=FALSE)
mapview::mapview(basin, alpha.regions=.08) + mapview::mapview(flowline_nldi, zcol = "PCTIMP2011CAT", legend = TRUE)
```

### Working with NARS data
In this example we demonstrate a data 'mashup' by grabbing NRSA data from the EPA National Aquatic Resource Surveys (NARS) website directly in R, pull particular StreamCat metrics for sites using `sc_get_data`, and compare landscape metrics with other NRSA metrics
```{r nars, results='hide'}
UA <- paste('Mozilla/5.0 (Windows NT 10.0; Win64; x64; rv:98.0)',
                    'Gecko/20100101 Firefox/98.0')
nrsa <- (data.table::fread(httr::content(httr::GET("https://www.epa.gov/sites/production/files/2015-09/siteinfo_0.csv", httr::add_headers(`User-Agent` = UA)), encoding = "UTF-8", as = "text"),stringsAsFactors = FALSE, data.table = F))

dplyr::glimpse(nrsa)

# Promote data frame to sf spatial points data frame
nrsa_sf <- sf::st_as_sf(nrsa, coords = c("LON_DD83", "LAT_DD83"), crs = 4269)

# Get COMIDs using nhdplusTools package
# nrsa$COMID<- NA
# for (i in 1:nrow(nrsa_sf)){
#   print (i)
#   nrsa_sf[i,'COMID'] <- discover_nhdplus_id(nrsa_sf[i,c('geometry')])
# }
load(system.file("extdata", "sample_nrsa_data.rda", package="StreamCatTools"))

# get particular StreamCat data for all these NRSA sites
# nrsa_sf$COMID <- as.character(nrsa_sf$COMID)
comids <- nrsa_sf$COMID
comids <- comids[!is.na(comids)]
comids <- comids[c(1:700)]
comids <- paste(comids,collapse=',')
df <- sc_get_data(metric='PctCrop2006', aoi='watershed', comid=comids)

# glimpse(df)
df$COMID <- as.integer(df$COMID)
nrsa_sf <- dplyr::left_join(nrsa_sf, df, by='COMID')
```

```{r nars_ggplot, warning=FALSE}
# download mmi from NARS web page
library(dplyr)
library(ggplot2)
mmi <- (data.table::fread(httr::content(httr::GET("https://www.epa.gov/sites/production/files/2015-09/bentcond.csv", httr::add_headers(`User-Agent` = UA)), encoding = "UTF-8", as = "text"),stringsAsFactors = FALSE, data.table = F))
# dplyr::glimpse(mmi)

# join mmi to NARS info data frame with StreamCat PctCrop metric
nrsa_sf <- dplyr::left_join(nrsa_sf, mmi[,c('SITE_ID','BENT_MMI_COND')], by='SITE_ID')
bxplt <- nrsa_sf %>% 
  tidyr::drop_na(BENT_MMI_COND) %>%
  ggplot2::ggplot(aes(x=PCTCROP2006WS, y=BENT_MMI_COND))+
  ggplot2::geom_boxplot()
suppressWarnings(print(bxplt))
```
