---
title: "Getting Started"
author: "Marc Weber"
output: 
  html_document:
    theme: flatly
    keep_md: yes
    number_sections: true
    highlighted: default 
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: no
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Getting Started With StreamCatTools StreamCatTools}
  %\VignetteEncoding{UTF-8}{inputenc}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  warning = FALSE, 
  message = FALSE
)
```

## Installing and loading StreamCatTools
To install, currently you need to install from GitHub using devtools
```{r, include=FALSE, message=FALSE, eval=FALSE}
library(devtools)
install_github('USEPA/StreamCatTools')
```

After installing load the library
```{r, include=FALSE, message=FALSE}
library(StreamCatTools)
```

## Background
The `StreamCatTools` package was designed to simplify the use of [StreamCat](https://www.epa.gov/national-aquatic-resource-surveys/streamcat-dataset) data in R, leveraging the new [API for StreamCat](https://intranet.ord.epa.gov/sswr/streamcat-web-tools). 

## StreamCat API
We can actually pull data into R from the [StreamCat API](https://java.epa.gov/StreamCAT/metrics/about.jsp) by simply using the `read_csv` function from the `readr` package.  We have to hard-wire parameters and are limited in the number of records returned through a `GET` request.
```{r API}
df <- readr::read_csv("https://java.epa.gov/StreamCAT/metrics?name=fert&areaOfInterest=catchment&comid=179", show_col_types = FALSE)
knitr::kable(df)
```

## List API parameters
List StreamCat parameters: Get a list of available StreamCat values for certain parameters using the `sc_get_params` function (right now just metric names and areas of interest for this function) via the API
```{r params}
region_params <- sc_get_params(param='areaOfInterest')

name_params <- sc_get_params(param='name')

print(paste0('region parameters are: ', paste(region_params,collapse = ', ')))
print(paste0('A selection of available StreamCat metrics include: ',paste(name_params[1:10],collapse = ', ')))
```

Look up the display name for a metric using the `sc_fullname` function via the API
```{r Example_One_fullnames}
metric='pcthbwet2011'
fullname <- sc_fullname(metric)
print(paste0('The full name for ',metric, ' is: ', paste(fullname,collapse = ', ')))
```

## Get data for COMIDs
In this example we access several variables, for several areas of interest, and for several COMIDs using the `sc_get_data` function.  Loads data into a tibble we can view.
```{r get_data}
df <- sc_get_data(metric='PctUrbMd2006,DamDens,TRIDens', aoi='riparian_catchment,catchment,watershed', comid='179,1337,1337420')
knitr::kable(df)
```

## Get data for county
In this example we access a couple variables at the watershed scale for the area of interest of a county (Benton County in this case) using the `sc_get_data` function. 
```{r counties}
df <- sc_get_data(metric='pctwdwet2006', aoi='watershed', county='41003')
knitr::kable(head(df))
```

## Get NLCD data
In this example we access National Land Cover Dataset (NLCD) data for 2001, just at the catchment level for several COMIDs using the `sc_nlcd` function.  Loads data into a tibble we can view.
```{r get_nlcd}
df <- sc_nlcd(year='2001', aoi='catchment',
              comid='179,1337,1337420')
knitr::kable(df)
```

We can also pass a couple years for a different area of interest for another region like a county.
```{r get_nlcd2}
df <- sc_nlcd(year='2006, 2019', aoi='watershed',
              county='41003')
knitr::kable(head(df))
```
## Get COMIDs
In this example we use the `sc_get_comid` function to find COMIDs for USGS stream gages we load into R. We use a .csv file with coordinate columns and a known coordinate reference system.  We'll use an example dataset of large dams in Oregon from the [Oregon Spatial Data Library]
```{r comids, eval=FALSE}
gages = readr::read_csv(system.file("extdata","Gages_flowdata.csv", package = "StreamCatTools"),show_col_types = FALSE)
# we'll just grab a few records to keep things simple
gages <- gages[1:10,c('SOURCE_FEA','STATION_NM','LON_SITE','LAT_SITE')]
gages_coms <- sc_get_comid(gages, xcoord='LON_SITE',                   ycoord='LAT_SITE', crsys=4269)

# Add the COMID we found back to gages data frame
gages$COMID <- strsplit(gages_coms, ",")[[1]]
df <- sc_get_data(metric='huden2010', aoi='watershed', comid=gages_coms)
df$COMID <- as.character(df$COMID)
gages <- dplyr::left_join(gages, df, by='COMID')
knitr::kable(head(gages))
```

## Get data for a hydroregion
In this example we access a couple watershed-only metrics for a particular NHDPlus hydroregion using the `sc_get_data` function.
```{r hydroregion}
df <- sc_get_data(metric='PctWdWet2006', aoi='watershed', region='17')
knitr::kable(head(df))
```

## Get data for CONUS
In this example we access a metric for conus using the `sc_get_data` function - this is shown for demonstration but not run as it takes a bit of time
```{r conus, eval=FALSE}
# df <- sc_get_data(metric='om', aoi='watershed', conus='true')
# knitr::kable(head(df))
```

