---
title: "Introduction"
author: "Marc Weber"
output: 
  html_document:
    theme: flatly
    keep_md: yes
    number_sections: true
    highlighted: default 
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: no
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Introduction to StreamCatTools}
  %\VignetteEncoding{UTF-8}{inputenc}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  warning = FALSE, 
  message = FALSE
)
```

## Installing and loading StreamCatTools
To install, currently you need to install from GitHub using devtools
```{r, include=FALSE, message=FALSE, eval=FALSE}
library(devtools)
install_github('USEPA/StreamCatTools')
```

After installing load the library
```{r, include=FALSE, message=FALSE}
library(StreamCatTools)
```

## Background
The `StreamCatTools` package was designed to simplify the use of [StreamCat](https://www.epa.gov/national-aquatic-resource-surveys/streamcat-dataset) data in R, leveraging the new [API for StreamCat](https://intranet.ord.epa.gov/sswr/streamcat-web-tools). 

### StreamCat API
We can actually pull data into R from the [StreamCat API](https://java.epa.gov/StreamCAT/metrics/about.jsp) by simply using the `read_csv` function from the `readr` package.  We have to hard-wire parameters and are limited in the number of records returned through a `GET` request.
```{r API}
df <- readr::read_csv("https://java.epa.gov/StreamCAT/metrics?name=fert&areaOfInterest=catchment&comid=179", show_col_types = FALSE)
knitr::kable(df)
```

### Examples of package functions
## Example One
List StreamCat parameters: Get a list of available StreamCat values for certain parameters using the `sc_get_params` function (right now just metric names and areas of interest for this function) via the API
```{r Example One}
region_params <- sc_get_params(param='areaOfInterest')

name_params <- sc_get_params(param='name')

print(paste0('region parameters are: ', paste(region_params,collapse = ', ')))
print(paste0('A selection of available StreamCat metrics include: ',paste(name_params[1:10],collapse = ', ')))
```

Look up the display name for a metric using the `sc_fullname` function via the API
```{r Example One fullnames}
metric='pcthbwet2011'
fullname <- sc_fullname(metric)
print(paste0('The full name for ',metric, ' is: ', paste(fullname,collapse = ', ')))
```

## Example Two
In this example we access several variables, for several areas of interest, and for several COMIDs using the `sc_get_data` function.  Loads data into a tibble we can view.
```{r Example Two}
df <- sc_get_data(metric='PctUrbMd2006,DamDens,TRIDens', aoi='riparian_catchment,catchment,watershed', comid='179,1337,1337420')
knitr::kable(df)
```

## Example Three
In this example we access a couple variables at the watershed scale for the area of interest of a county (Benton County in this case) using the `sc_get_data` function. 
```{r Example Three}
df <- sc_get_data(metric='pctwdwet2006', aoi='watershed', county='41003')
knitr::kable(head(df))
```

## Example Four
In this example we use the `sc_get_comid` function to find COMIDs for USGS stream gages we load into R. We use a .csv file with coordinate columns and a known coordinate reference system.  We'll use an example dataset of large dams in Oregon from the [Oregon Spatial Data Library]
```{r Example Four}
gages = readr::read_csv(system.file("extdata/Gages_flowdata.csv", package = "StreamCatTools"),show_col_types = FALSE)
# we'll just grab a few records to keep things simple
gages <- gages[1:10,c('SOURCE_FEA','STATION_NM','LON_SITE','LAT_SITE')]
gages_coms <- sc_get_comid(gages, xcoord='LON_SITE',                   ycoord='LAT_SITE', crsys=4269)

# Add the COMID we found back to gages data frame
gages$COMID <- strsplit(gages_coms, ",")[[1]]
df <- sc_get_data(metric='huden2010', aoi='watershed', comid=gages_coms)
df$COMID <- as.character(df$COMID)
gages <- dplyr::left_join(gages, df, by='COMID')
knitr::kable(head(gages))
```

## Example Five
In this example we access a couple watershed-only metrics for a particular NHDPlus hydroregion using the `sc_get_data` function. 
```{r Example Five}
df <- sc_get_data(metric='PctWdWet2006', aoi='watershed', region='17')
knitr::kable(head(df))
```

## Example Six
In this example we access a metric for conus using the `sc_get_data` function - this is shown for demonstration but not run as it takes a bit of time
```{r Example Six, eval=FALSE}
# df <- sc_get_data(metric='om', aoi='watershed', conus='true')
# knitr::kable(head(df))
```

## Example Seven
In this example we access a single variable for the Calapooia River using `sc_get_data` function. We then use the `nhdplusTools` library to grab flowlines and watershed for the Calapooia, plot the selected StreamCat metric for the Calapooia River and show the watershed.
```{r Example Seven, results='hide'}
start_comid = 23763529
nldi_feature <- list(featureSource = "comid", featureID = start_comid)

flowline_nldi <- nhdplusTools::navigate_nldi(nldi_feature, mode = "UT", data_source = "flowlines", distance=5000)

# get StreamCat metrics
temp_dir <- '.'
nhdplus <- nhdplusTools::subset_nhdplus(comids = as.integer(flowline_nldi$UT_flowlines$nhdplus_comid), output_file = file.path(temp_dir, "nhdplus.gpkg"),nhdplus_data = "download",overwrite = TRUE,flowline_only = FALSE)

names(nhdplus)
cats <- nhdplus$CatchmentSP
comids <- paste(cats$featureid,collapse=",",sep="")

df <- sc_get_data(metric='PctImp2011', aoi='catchment', comid=comids)

flowline_nldi <- flowline_nldi$UT_flowlines
flowline_nldi$PCTIMP2011CAT <- df$PCTIMP2011CAT[match(flowline_nldi$nhdplus_comid, df$COMID)]

basin <- nhdplusTools::get_nldi_basin(nldi_feature = nldi_feature)
```

```{r Example Seven Pt2}
library(mapview)
mapview::mapviewOptions(fgb=FALSE)
mapview::mapview(basin, alpha.regions=.08) + mapview::mapview(flowline_nldi, zcol = "PCTIMP2011CAT", legend = TRUE)
```

## Example Eight
In this example we demonstrate a data 'mashup' by grabbing NRSA data from the EPA National Aquatic Resource Surveys (NARS) website directly in R, pull particular StreamCat metrics for sites using `sc_get_data`, and compare landscape metrics with other NRSA metrics
```{r Example Eight, results='hide'}
nrsa <- readr::read_csv('https://www.epa.gov/sites/production/files/2015-09/siteinfo_0.csv')
# dplyr::glimpse(nrsa)

# Promote data frame to sf spatial points data frame
nrsa_sf <- sf::st_as_sf(nrsa, coords = c("LON_DD83", "LAT_DD83"), crs = 4269)

# Get COMIDs using nhdplusTools package
# nrsa$COMID<- NA
# for (i in 1:nrow(nrsa_sf)){
#   print (i)
#   nrsa_sf[i,'COMID'] <- discover_nhdplus_id(nrsa_sf[i,c('geometry')])
# }
load(system.file("extdata", "sample_nrsa_data.rda", package="StreamCatTools"))

# get particular StreamCat data for all these NRSA sites
# nrsa_sf$COMID <- as.character(nrsa_sf$COMID)
comids <- nrsa_sf$COMID
comids <- comids[!is.na(comids)]
comids <- comids[c(1:925)]
comids <- paste(comids,collapse=',')
df <- sc_get_data(metric='PctCrop2006', aoi='watershed', comid=comids)

# glimpse(df)
df$COMID <- as.integer(df$COMID)
nrsa_sf <- dplyr::left_join(nrsa_sf, df, by='COMID')
```

```{r Example Eight ggplot}
# download mmi from NARS web page
library(dplyr)
library(ggplot2)
mmi <- readr::read_csv('https://www.epa.gov/sites/production/files/2015-09/bentcond.csv',show_col_types = FALSE)
dplyr::glimpse(mmi)

# join mmi to NARS info data frame with StreamCat PctCrop metric
nrsa_sf <- dplyr::left_join(nrsa_sf, mmi[,c('SITE_ID','BENT_MMI_COND')], by='SITE_ID')
bxplt <- nrsa_sf %>% 
  tidyr::drop_na(BENT_MMI_COND) %>%
  ggplot(aes(x=PCTCROP2006WS, y=BENT_MMI_COND))+
  geom_boxplot()
suppressWarnings(print(bxplt))
```