---
title: "Working with LakeCat data"
output: 
  html_document:
    theme: flatly
    keep_md: yes
    number_sections: true
    highlighted: default 
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: no
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Working with LakeCat data}
  %\VignetteEncoding{UTF-8}{inputenc}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  warning = FALSE, 
  message = FALSE
)
```

## Accessing and working with [LakeCat](https://www.epa.gov/national-aquatic-resource-surveys/lakecat-dataset)
The `StreamCatTools` package was designed to simplify the use of [StreamCat](https://www.epa.gov/national-aquatic-resource-surveys/streamcat-dataset) data in R, leveraging the new [API for StreamCat](https://intranet.ord.epa.gov/sswr/streamcat-web-tools).  We now have functionality in the [StreamCat API](https://java.epa.gov/StreamCAT/metrics/about.jsp) for accessing and working with LakeCat data and have added functions in `StreamCatTools` to access LakeCata data in R using the API.

We can actually pull data into R from LakeCat by simply using the `read_csv` function from the `readr` package. We have to hard-wire parameters and are limited in the number of records returned through a `GET` request.
```{r, message=FALSE, eval=FALSE}
df <- readr::read_csv("https://java.epa.gov/StreamCAT/LakeCat/metrics?name=PctUrbMd2006&areaOfInterest=catchment&comid=22538788", show_col_types = FALSE)
```

## List LakeCat API parameters

List LakeCat parameters: Get a list of available LakeCat values for certain parameters using the `lc_get_params` function (right now just metric names and areas of interest for this function) via the API

```{r params}
library(StreamCatTools)
region_params <- lc_get_params(param='areaOfInterest')

name_params <- lc_get_params(param='name')

print(paste0('region parameters are: ', paste(region_params,collapse = ', ')))
print(paste0('A selection of available LakrCat metrics include: ',paste(name_params[1:10],collapse = ', ')))
```

Look up the display name or names for a metric using the `lc_fullname` function via the API
```{r fullnames}
metric='pcthbwet2011'
fullname <- lc_fullname(metric)
fullname
```

```{r fullnames2}
metric='pctdecid2019,fert'
fullname <- lc_fullname(metric)
fullname
```

## Get Waterbody COMIDs
In this example we use the `lc_get_comid` function to find COMIDs for a set of example lake locations we load into R.`lc_get_comid` is just a simple wrapper for `get_waterbodies` in the [nhdplusTools](https://doi-usgs.github.io/nhdplusTools/) R package. We can then use the COMIDs we derive for our lake locations to get LakeCat metrics for these lakes as we show in after this.
```{r comids, warning=FALSE, message=FALSE}
dd <- data.frame(x = c(-89.198,-114.125,-122.044),
                 y = c(45.502,47.877,43.730)) |> 
  sf::st_as_sf(coords = c('x', 'y'), crs = 4326)
    
comids <- lc_get_comid(dd)
```

## Get data for COMIDs

In this example we access several variables, for several areas of interest, and for several COMIDs using the `lc_get_data` function. We'll show using both the COMIDS we derived with `lc_get_comid` function in previous chunk as well as with known COMIDS for NHDPlus waterbodies. Loads data into a tibble we can view.

```{r get_data}
df <- lc_get_data(metric='PctUrbMd2006,DamDens', aoi='catchment,watershed', comid=comids)
knitr::kable(df)
```

```{r get_data2}
df <- lc_get_data(metric='PctUrbMd2006,DamDens', aoi='catchment,watershed', comid='23783629,23794487,23812618')
knitr::kable(df)
```

## Get NLCD data

In this example we access National Land Cover Dataset (NLCD) data for 2019, just at the catchment level for several COMIDs using the `lc_nlcd` function. Loads data into a tibble we can view.

```{r get_nlcd}
df <- lc_nlcd(comid='23783629,23794487,23812618', year='2019', aoi='watershed')
knitr::kable(df)
```
