---
title: "National Nutrient Inventory (NNI) Data in StreamCat"
output: 
  html_document:
    theme: flatly
    keep_md: yes
    number_sections: true
    highlighted: default 
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: no
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Applications withs StreamCatTools}
  %\VignetteEncoding{UTF-8}{inputenc}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  warning = FALSE, 
  message = FALSE
)
```

## Access Farm Nitrogen data for a watershed
In this example we access a single National Nutrient Inventory (NNI) metric for the Calapooia River basin using the `sc_get_data` function. We use the `nhdplusTools` library to pull in flowlines and the watershed boundary for the Calapooia River basin, plot the selected NNI metric for the Calapooia River and show the watershed.
```{r farmn, results='hide'}
library(StreamCatTools)
start_comid = 23763517
nldi_feature <- list(featureSource = "comid", featureID = start_comid)

flowline_nldi <- nhdplusTools::navigate_nldi(nldi_feature, mode = "UT", data_source = "flowlines", distance=5000)

# get StreamCat metrics
comids <- paste(as.integer(flowline_nldi$UT_flowlines$nhdplus_comid), collapse=",",sep="")

df <- sc_get_data(metric='n_ff_2016', aoi='cat', comid=comids, showAreaSqKm=TRUE)

flowline_nldi <- flowline_nldi$UT_flowlines
flowline_nldi$Farm_Nitrogon_2016 <- df$n_ff_2016cat[match(flowline_nldi$nhdplus_comid, df$comid)]

basin <- nhdplusTools::get_nldi_basin(nldi_feature = nldi_feature)
```

## Map the Results
```{r map1, results='hide', message='false'}
library(ggplot2)
library(ggspatial)
flowline_nldi |> 
  ggplot() + geom_sf(aes(colour = Farm_Nitrogon_2016)) + 
  scale_y_continuous() +
  scale_color_distiller(palette = "Spectral") +
  labs(color = "Kg Nitrogen") + 
  theme_minimal(12) + 
  ggtitle('Farm Nitrogen at the Catchment Scale for \nthe Calapooia River Basin 2016')
```

## Look at change through time Calapooia Farm Nitrogen
```{r change, results='hide', message='false'}
df1 <- sc_get_data(metric='n_ff_1987', aoi='cat', comid=comids)
df2 <- sc_get_data(metric='n_ff_2017', aoi='cat', comid=comids)
df2$n_ff_1987cat <- df1$n_ff_1987cat[match(df2$comid, df1$comid)]
df2$Farm_Nitrogen_Difference <- df2$n_ff_2017cat-df2$n_ff_1987cat
flowline_nldi$Farm_Nitrogen_Difference <- df2$Farm_Nitrogen_Difference[match(flowline_nldi$nhdplus_comid, df2$comid)]
```

## Map the Results
```{r map2, results='hide', message='false'}
flowline_nldi |> 
  ggplot() + geom_sf(aes(colour = Farm_Nitrogen_Difference)) + 
  scale_y_continuous() + 
  labs(color = "Kg Nitrogen")+ 
  scale_color_distiller(palette = "Spectral") +
  theme_minimal(12) + 
  ggtitle('Farm Nitrogen Change at the Catchment Scale for \nthe Calapooia River Basin from 1987 to 2017')
```

## Crop Fixation for the Calapooia Watershed
```{r crop, results='hide', message='false'}
options(scipen=3)
# get StreamCat metrics
df <- sc_get_data(metric='n_cf_2017', aoi='ws', comid=comids)

flowline_nldi$CropNFixation <- df$n_cf_2017ws[match(flowline_nldi$nhdplus_comid, df$comid)]
```

## Map the Results
```{r map3, results='hide', message='false'}
flowline_nldi |> 
  ggplot() + geom_sf(aes(colour = CropNFixation)) + 
  scale_y_continuous() +
  scale_color_distiller(palette = "Spectral") +
  labs(colour = "Kg Nitrogen") + 
  theme_minimal(12) + 
  ggtitle('Watershed Crop Nitrogen Fixation \nfor the Calapooia River Basin 2017')
```