---
title: "Start Here"
output: 
  html_document:
    theme: flatly
    keep_md: yes
    number_sections: true
    highlighted: default 
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: no
    toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Start Here}
  %\VignetteEncoding{UTF-8}{inputenc}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  warning = FALSE, 
  message = FALSE
)
```

## Installing and loading StreamCatTools
To install, you can install from CRAN:
```{r, message=FALSE, eval=FALSE}
install.packages("StreamCatTools")
```

Or you can install from GitHub:
```{r, message=FALSE, eval=FALSE}
library(devtools)
install_github('USEPA/StreamCatTools')
```

After installing load the library
```{r, message=FALSE}
library(StreamCatTools)
```

## Background
The `StreamCatTools` package was designed to simplify the use of [StreamCat](https://www.epa.gov/national-aquatic-resource-surveys/streamcat-dataset) data in R, leveraging the new [API for StreamCat](https://api.epa.gov/StreamCat/streams/metrics). 

# Vignettes in `StreamCatTools`

There are three additional vignettes in `StreamCatTools` that are available on our website [here](https://usepa.github.io/StreamCatTools/) under the "Articles" tab:

1. Introduction
  * An overview of the basic fucntionality of the package
2. LakeCat
  * An overview of accessing data for lakes within the package and functions to do so
3. Applications
  * A more in depth walk through with specific use-case examples and demonstrations

These vignettes cover some of the core functions (and arguments within those functions) in `StreamCatTools`. Help files for a particular function are viewable by running `?function_name` after loading `StreamCatTools`. For example, to learn more about the `sc_get_data()` function, run `?sc-get_data`.

# Citation information

If you use `StreamCatTools` in your work, please cite it. You can view the most recent citation by running
```{r}
citation("StreamCatTools")
```
