---
title: "Introduction to the StreamCatTools Package"
author: "Marc Weber"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    keep_md: yes
    theme: yeti
    highlighted: default 
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
      number_sections: true
vignette: >
  %\VignetteIndexEntry{Introduction to StreamCatTools}
  %\VignetteEncoding{UTF-8}{inputenc}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: inline
---


```{r setup, include=FALSE, message=FALSE}
# requires devtools to install
# install.packages('devtools')
library(devtools)

# install from repository
# install_github('USEPA/StreamCatTools')
library(StreamCatTools)
library(sf)
library(nhdplusTools)
library(readr)
library(mapview)
library(dplyr)
library(knitr)
library(readr)
library(purrr)
library(tidyr)
library(ggplot2)
library(jsonlite)
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      error=FALSE,
                      comment = "#>",
                      fig.height = 7,
                      fig.width = 7)
sessionInfo()
```

The `StreamCatTools` package was designed to simplify the use of [StreamCat](https://www.epa.gov/national-aquatic-resource-surveys/streamcat-dataset-0) data in R, leveraging the new [API for StreamCat](https://intranet.ord.epa.gov/sswr/streamcat-web-tools). 

## StreamCat API
We can actually pull data into R from the [StreamCat API](http://v26267mcpk506/StreamCat/v1/stable/metrics/about.jsp) by simply using `read_csv` function from the `readr` package.  We have to hard-wire parameters and are limited in the number of records returned through a `GET` request.
```{r API}
df <- read_csv("https://v26267mcpk506/StreamCat/v1/stable/metrics?name=fert&areaOfInterest=watershed&comid=179")
kable(df)
```

## Example One
List parameters -get a list of available StreamCat values for certain parameters using the `sc_get_params` function (right now just metric names and areas of interest for this function) via the API
```{r Example One}
region_params <- sc_get_params(param='areaOfInterest')

name_params <- sc_get_params(param='name')

print(paste0('region parameters are: ', paste(region_params,collapse = ', ')))
print(paste0('A selection of available StreamCat metrics include: ',paste(name_params[1:10],collapse = ', ')))
```

Look up the display name for a metric using the `sc_fullname` function via the API
```{r Example One fullnames}
metric='pcthbwet2011'
fullname <- sc_fullname(metric)
print(paste0('The full name for ',metric, ' is: ', paste(fullname,collapse = ', ')))
```

## Example Two
Access several variables for several areas of interest and a couple COMIDs using the `sc_get_data` function.  Loads data into a tibble we can view.
```{r Example Two}
df <- sc_get_data(metric='PctUrbMd2006,DamDens,TRIDens', aoi='riparian_catchment,catchment,watershed', comid='179,1337,1337420')
kable(df)
```

## Example Three
Access a couple watershed-only variables for the area of interest of a county (Benton County in this case) using the `sc_get_data` function. 
```{r Example Three}
df <- sc_get_data(metric='pctwdwet2006', aoi='watershed', county='41003')
kable(head(df))
```

## Example Four
Use the `sc_get_comid` function to find COMIDs for USGS stream gages we load into R. We use a .csv file with coordinate columns and a known coordinate reference system.  We'll use an example dataset of large dams in Oregon from the [Oregon Spatial Data Library]
```{r Example Four}
gages = read_csv(system.file("extdata/Gages_flowdata.csv", package = "StreamCatTools"))
# we'll just grab a few records to keep things simple
dam_coms <- sc_get_comid(gages[1:10,], xcoord='LON_SITE', 
ycoord='LAT_SITE', crdsys=4269)
df <- sc_get_data(metric='huden2010', aoi='watershed', comid=dam_coms)
kable(head(df))
```

## Example Five
Access a couple watershed-only metrics for a particular hydroregion using the `sc_get_data` function. 
```{r Example Five}
df <- sc_get_data(metric='PctWdWet2006', aoi='watershed', region='17')
kable(head(df))
```

## Example Six
Access a metric for conus using the `sc_get_data` function.
```{r Example Six}
df <- sc_get_data(metric='om', aoi='watershed', conus='true')
kable(head(df))
```

## Example Seven
Access a single variable for the Calapooia River using `sc_get_data` function. Use `nhdplusTools` library to grab flowlines and watershed for the Calapooia, plot selected StreamCat metric for Calapooia and show the watershed.
```{r Example Seven, results='hide'}
start_comid = 23763529
nldi_feature <- list(featureSource = "comid", featureID = start_comid)

flowline_nldi <- navigate_nldi(nldi_feature, mode = "UT", data_source = "flowlines", distance=5000)

# get StreamCat metrics
temp_dir <- '.'
nhdplus <- subset_nhdplus(comids = flowline_nldi$UT_flowlines$nhdplus_comid, output_file = file.path(temp_dir, "nhdplus.gpkg"),nhdplus_data = "download",overwrite = TRUE,flowline_only = FALSE)

names(nhdplus)
cats <- nhdplus$CatchmentSP
comids <- paste(cats$featureid,collapse=",",sep="")

df <- sc_get_data(metric='PctImp2011', aoi='catchment', comid=comids)

flowline_nldi <- flowline_nldi$UT_flowlines
flowline_nldi$PCTIMP2011CAT <- df$PCTIMP2011CAT[match(flowline_nldi$nhdplus_comid, df$COMID)]

basin <- get_nldi_basin(nldi_feature = nldi_feature)
```

```{r Example Seven Pt2}
# mapview(basin, alpha.regions=.08) + mapview(flowline_nldi, zcol = "PCTIMP2011CAT", legend = TRUE)
mapview(flowline_nldi, zcol = "PCTIMP2011CAT", legend = TRUE)
```

## Example Eight
Grab NRSA data from NARS website directly in R, pull particular StreamCat metrics for sites using `sc_get_data`, and compare landscape metrics with other NRSA metrics
```{r Example Eight, results='hide'}
nrsa <- read_csv('https://www.epa.gov/sites/production/files/2015-09/siteinfo_0.csv')
glimpse(nrsa)

# Promote data frame to sf spatial points data frame
nrsa_sf <- st_as_sf(nrsa, coords = c("LON_DD83", "LAT_DD83"), crs = 4269)

# Get COMIDs using nhdplusTools package
# nrsa$COMID<- NA
# for (i in 1:nrow(nrsa_sf)){
#   print (i)
#   nrsa_sf[i,'COMID'] <- discover_nhdplus_id(nrsa_sf[i,c('geometry')])
# }
load(system.file("extdata", "sample_nrsa_data.rda", package="StreamCatTools"))

# get particular StreamCat data for all these NRSA sites
# nrsa_sf$COMID <- as.character(nrsa_sf$COMID)
comids <- nrsa_sf$COMID
comids <- comids[!is.na(comids)]
comids <- comids[c(1:925)]
comids <- paste(comids,collapse=',')
df <- sc_get_data(metric='PctCrop2006', aoi='watershed', comid=comids)

glimpse(df)
df$COMID <- as.integer(df$COMID)
nrsa_sf <- left_join(nrsa_sf, df, by='COMID')
```

```{r Example Eight ggplot}
# download mmi from NARS web page
mmi <- read_csv('https://www.epa.gov/sites/production/files/2015-09/bentcond.csv')
glimpse(mmi)

# join mmi to NARS info data frame with StreamCat PctCrop metric
nrsa_sf <- left_join(nrsa_sf, mmi[,c('SITE_ID','BENT_MMI_COND')], by='SITE_ID')
nrsa_sf %>% 
  drop_na(BENT_MMI_COND) %>%
  ggplot(aes(x=PCTCROP2006WS, y=BENT_MMI_COND))+
  geom_boxplot()
```