---
title: "SnapSitesToNHDPlus"
author: "Marc Weber"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    theme: yeti
    highlighted: default 
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Read in point location data

```{r sites, warning=FALSE, error=FALSE, message=FALSE}
library(dplyr); 
library(ggplot2); 
library(sf); 
library(readr)
library(mapview)
library(leaflet)
library(foreign)
library(knitr)

df_locs <- read_csv('VA_FF_WVA_MD_Other_envData.csv')
# head(as.data.frame(df_locs))
df_locs <- st_as_sf(df_locs, 
                    coords = c("LongitudeDD", "LatitudeDD"), # for point data
                    remove = F, # don't remove these lat/lon cols from df
                    crs = 4269) # add projection 
```

## Assign NHDPlus regions to sites
Project both df_locs and flowlines to Albers, asign sites to NHDPlus regions, and write out temp NHD lines
```{r NHDPlus, eval=FALSE}
regions <- st_read('H:/NHDPlusV21/NHDPlusGlobalData/BoundaryUnit.shp')
regions <- regions %>% 
  dplyr::filter(UnitType=='VPU') %>%
  dplyr::select(DrainageID, UnitID, geometry)

df_locs <- df_locs %>% dplyr::select(UID, StationID, LongitudeDD, LatitudeDD, state)
regions <- st_transform(regions, st_crs(df_locs))
df_locs <- st_join(df_locs, regions) 

df_locs$UnitID <- droplevels(df_locs$UnitID)

i=0
for (l in levels(df_locs$UnitID)){
  d <- regions$DrainageID[regions$UnitID == l]
  if (i==0){
    nhd <- st_read(paste0('H:/NHDPlusV21/NHDPlus',d,'/NHDPlus',l,'/NHDSnapshot/Hydrography/NHDFlowline.shp'))
    nhd <- st_transform(nhd, st_crs(df_locs))
    nhd <- st_zm(nhd, drop = TRUE, what = "ZM")
    nhd <- st_crop(nhd, st_bbox(df_locs))
    nhd2 <- st_read(paste0('H:/NHDPlusV21/NHDPlus',d,'/NHDPlus',l,'/NHDSnapshot/Hydrography/NHDFlowline.shp'))
    nhd2 <- nhd2 %>% 
      dplyr::filter(COMID %in% nhd$COMID)
  }
  if (i > 0){
    temp <- st_read(paste0('H:/NHDPlusV21/NHDPlus',d,'/NHDPlus',l,'/NHDSnapshot/Hydrography/NHDFlowline.shp'))
    temp <- st_transform(temp, st_crs(df_locs))
    temp <- st_zm(temp, drop = TRUE, what = "ZM")
    temp <- st_crop(temp, st_bbox(df_locs))
    temp2 <- st_read(paste0('H:/NHDPlusV21/NHDPlus',d,'/NHDPlus',l,'/NHDSnapshot/Hydrography/NHDFlowline.shp'))
    temp2 <- temp2 %>% 
      dplyr::filter(COMID %in% temp$COMID)
    nhd2 <- rbind(nhd2, temp2)
  }
  i = i+1
}

nhd2 <- nhd2 %>% 
  st_transform(crs = 102003) # convert to Albers Equal Area
df_locs <- df_locs %>% 
  st_transform(crs = 102003) # convert to Albers Equal Area


# Write out temporary projected shapefiles to use for linear referencing sites
st_write(df_locs, 'H:/WorkingData/Junk/sites_for_lr.shp')
st_write(nhd2, 'H:/WorkingData/Junk/nhd_for_lr.shp')
```

## Set up reticulate to run python from R
You may need to install reticulate, and then you need to set your python path (where your python excutable is located that you want to use), use a conda environment, or use a virtual environment for python
```{r reticulate, eval=FALSE}
# install.packages("reticulate")
library(reticulate)
# use_python("C:/Python27/ArcGIS10.6") # A typical installation example...
use_condaenv("StreamCat") # I use a conda environment installation of python
```

## Snap sites to flowlines
Use linear referencing with ArcPy and create route if streamlines are not already a route
```{python, eval=FALSE}
import arcpy
# If streamlines are not routes, you will need to modify Create Routes below to fit your needs
# Execute CreateRoutes
# arcpy.CreateRoutes_lr(in_lines, route_ID_field, out_routes, "LENGTH", "#", "#", "LOWER_LEFT", 0.00018939394)

# Execute LocateFeaturesAlongRoutes
pts = 'H:/WorkingData/Junk/sites_for_lr.shp'
nhd = 'H:/WorkingData/Junk/nhd_for_lr.shp'
props = "RID POINT MEAS"
arcpy.LocateFeaturesAlongRoutes_lr(pts, nhd, "REACHCODE", "3000 Meters", "H:/WorkingData/Junk/LR_Sites.dbf", props, "FIRST", "DISTANCE", "ZERO", "FIELDS", "M_DIRECTON")
```

### View locations and NHDPlus
```{r mapview, eval=TRUE}
df_locs <- st_read('H:/WorkingData/Junk/sites_for_lr.shp')
m <- mapview(df_locs)
m@map = m@map %>% addWMSTiles(group = 'NHDPlus',
"https://watersgeo.epa.gov/arcgis/services/NHDPlus_NP21/NHDSnapshot_NP21/MapServer/WmsServer?",
layers  = 4,
options = WMSTileOptions(format = "image/png", transparent = TRUE),
attribution = "") %>% mapview:::mapViewLayersControl(names = c("NHDPlus"))
m
```


### Flag results
Check for sites that were far from streamlines or that had multiple streamlines within a given buffer distance of original site locations. Note negative values indicate direction from line.  I used example values of > 20 meters from stream or more than 2 streams within a 50 meter buffer of sites to be conservative.

```{r flags, eval=TRUE, warning=FALSE, error=FALSE, message=FALSE}
nhd <- st_read('H:/WorkingData/Junk/nhd_for_lr.shp')
sites_lr <- read.dbf('H:/WorkingData/Junk/LR_Sites.dbf')

a <- ggplot(sites_lr, aes(x = Distance))
a + geom_histogram(bins = 30, color = "black", fill = "gray") +
  geom_vline(aes(xintercept = mean(Distance)), 
             linetype = "dashed", size = 0.6)
summary(abs(sites_lr$Distance))

# Drop sites more than 20 meters from streamlines just as an example distance flag
too_far <- sites_lr[abs(sites_lr$Distance) > 20,]

# Back to Albers projection for creating buffers
# df_locs <- df_locs %>% 
#   st_transform(crs = 102003) # convert to Albers Equal Area
bufs <- st_buffer(df_locs, dist=50)
nhd <- st_zm(nhd, drop = TRUE, what = "ZM")
bufs <- st_intersection(bufs, nhd)
too_many_streams <- bufs %>%
  group_by(UID) %>%
  filter(n_distinct(COMID) > 1)

# Add flag
df_locs$Flag <- ifelse((df_locs$UID %in% too_far$UID) | (df_locs$UID %in% too_many_streams$UID), 'Need to Check', 'OK')
kable(table(df_locs$Flag))
st_geometry(bufs) <- NULL 
df_locs <- left_join(df_locs, bufs[,c('UID','COMID','REACHCODE','GNIS_ID','GNIS_NAME')])
st_geometry(df_locs) <- NULL 
write_csv(df_locs, 'F:/Git Projects/LouReynolds/Sites_with_Flags_and_NHDPlusInfo.csv')
```

