---
title: "StreamCatTools Demo"
author: "Marc Weber"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    theme: yeti
    highlighted: default 
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: yes
---

## Install StreamCatTools
Also make sure you have nhdplusTools installed - I am leveraging some functions from that package to complement grabbing Streamcat data.  I'm also using mapview package.
```{r load libraries, eval=TRUE, warning=FALSE, message=FALSE}
# requires devtools to install
# install.packages('devtools')
library(devtools)

# install from repository
# install_github('USEPA/StreamCatTools')
library(StreamCatTools)
library(sf)
library(nhdplusTools)
library(readr)
library(mapview)
library(dplyr)
library(knitr)
library(readr)
library(purrr)
library(tidyr)
library(ggplot2)
library(jsonlite)
sessionInfo()
```

## Example One
Get a list of available StreamCat values for certain parameters using the `get_streamcat_params` function (right now just metric names and areas of interest for this function) via the API
```{r Example One, warning=FALSE, error=FALSE, message=FALSE}
area_params <- get_streamcat_params(param='area')
name_params <- get_streamcat_params(param='name')
print(paste0('Area of interest available parameters are: ', paste(area_params,collapse = ', ')))
print(paste0('A selection of available StreamCat metrics include: ',paste(name_params[1:10],collapse = ', ')))
```

## Example Two
Access several variables for several areas of interest and a couple COMIDs using theh `get_streamcat_data` function.  Loads data into a tibble we can view.
```{r Example Two, warning=FALSE, error=FALSE, message=FALSE}
df <- get_streamcat_data(metric='PctUrbMd2006,DamDens,TRIDens', aoi='riparian_catchment,catchment,watershed', comid='179,1337,1337420')
kable(df)
```

## Example Three
We can actually pull data into R from the StreamCat API by simply using `read_csv` function from the `readr` package.  We have to hard-wire paramaters and are limited in number of records returned through a `GET` request
```{r Example Three, warning=FALSE, error=FALSE, message=FALSE}
df <- read_csv("http://v26267mcpk506/StreamCat/v1/stable/metrics?name=fert&areaOfInterest=watershed&comid=179")
kable(df)
```

## Example Four
Access a couple watershed-only variables for a county (Benton County in this case) using the `get_streamcat_data` function. 
```{r Example Four, warning=FALSE, error=FALSE, message=FALSE}
df <- get_streamcat_data(metric='PctWdWet2006', aoi='watershed', county='41003')
kable(head(df))
```

## Example Five
Access a couple watershed-only metrics for a particular hydroregion using the `get_streamcat_data` function. 
```{r Example Five, warning=FALSE, error=FALSE, message=FALSE}
df <- get_streamcat_data(metric='PctWdWet2006', aoi='watershed', region='17')
kable(head(df))
```

## Example Six
Access a single variable for the Calapooia River using `get_streamcat_data` function. Use `nhdplusTools` library to grab flowlines and watershed for the Calapooia, plot selected StreamCat metric for Calapooia and show the watershed.
```{r Example Six, warning=FALSE, error=FALSE, message=FALSE, fig.width=9}
discover_nldi_sources()$source
start_comid = 23763529
nldi_feature <- list(featureSource = "comid", featureID = start_comid)
discover_nldi_navigation(nldi_feature)

flowline_nldi <- navigate_nldi(nldi_feature, mode = "upstreamTributaries", data_source = "")

# get StreamCat metrics
temp_dir <- 'C:/Users/mweber/temp'
nhdplus <- subset_nhdplus(comids = flowline_nldi$nhdplus_comid, output_file = file.path(temp_dir, "nhdplus.gpkg"),nhdplus_data = "download",overwrite = TRUE, return_data = FALSE)

st_layers(nhdplus)
cats <- read_sf(dsn=nhdplus, layer='CatchmentSP')
comids <- paste(cats$featureid,collapse=",",sep="")

df <- get_streamcat_data(metric='PctImp2011', aoi='catchment', comid=comids)

flowline_nldi$PCTIMP2011CAT <- df$PCTIMP2011CAT[match(flowline_nldi$nhdplus_comid, df$COMID)]

basin <- get_nldi_basin(nldi_feature = nldi_feature)

mapview(basin, alpha.regions=.08) + mapview(flowline_nldi, zcol = "PCTIMP2011CAT", legend = TRUE)
```

## Example Seven
Grab NRSA data from NARS website directly in R, pull particular StreamCat metrics for sites using `get_streamcat_data`, and compare landscape metrics with other NRSA metrics
```{r Example Seven, warning=FALSE, error=FALSE, message=FALSE}
nrsa <- read_csv('https://www.epa.gov/sites/production/files/2015-09/siteinfo_0.csv')
glimpse(nrsa)

# Promote data frame to sf spatial points data frame
nrsa_sf <- st_as_sf(nrsa, coords = c("LON_DD83", "LAT_DD83"), crs = 4269)

# Get COMIDs using nhdplusTools package
# nrsa$COMID<- NA
# for (i in 1:nrow(nrsa_sf)){
#   print (i)
#   nrsa_sf[i,'COMID'] <- discover_nhdplus_id(nrsa_sf[i,c('geometry')])
# }
load("L:/Public/mweber/example.RData")

# get particular StreamCat data for all these NRSA sites
# nrsa_sf$COMID <- as.character(nrsa_sf$COMID)
comids <- nrsa_sf$COMID
comids <- comids[!is.na(comids)]
comids <- comids[c(1:925)]
comids <- paste(comids,collapse=',')
df <- get_streamcat_data(metric='PctCrop2006', aoi='watershed', comid=comids)

glimpse(df)
df$COMID <- as.integer(df$COMID)
nrsa_sf <- left_join(nrsa_sf, df, by='COMID')
```

```{r Example Six ggplot, warning=FALSE, error=FALSE, message=FALSE, fig.width=9}
# download mmi from NARS web page
mmi <- read_csv('https://www.epa.gov/sites/production/files/2015-09/bentcond.csv')
glimpse(mmi)

# join mmi to NARS info data frame with StreamCat PctCrop metric
nrsa_sf <- left_join(nrsa_sf, mmi[,c('SITE_ID','BENT_MMI_COND')], by='SITE_ID')
nrsa_sf %>% 
  drop_na(BENT_MMI_COND) %>%
  ggplot(aes(x=PCTCROP2006WS, y=BENT_MMI_COND))+
  geom_boxplot()

```